|

### 2023-06-23 src/docs_paper/paper.Rmd
Update to new repo name

---
title: Understanding models for spatial structure in small-area estimation
classoption: ba
author:
  - firstname: Adam 
    lastname: Howes
    email: ath19@ic.ac.uk
    affiliationref: A,B
    footnoterefs: 
      - t1
  - firstname: Jeffrey W.
    lastname: Eaton
    affiliationref: C,B
  - firstname: Seth R.
    lastname: Flaxman
    affiliationref: D
affiliations:
  - ref: A
    name: Imperial College London 
    address: Department of Mathematics
  - ref: B
    name: Imperial College London
    address: MRC Centre for Global Infectious Disease Analysis
  - ref: C
    name: Harvard University
    address: Center for Communicable Disease Dynamics, Harvard T.H. Chan School of Public Health
  - ref: D
    name: University of Oxford
    address: Department of Computer Science 
footnotes:
  - ref: t1
    text: "Corresponding Author"
supplements:
  - title: Appendix
    description: All further materials for this study.
acknowledgements: |
  | AH thanks Robert Ashton for help with \textsf{R} package development, and Harrison Zhu, Theo Rashid, Tim Wolock, Samir Bhatt, Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for helpful conversations. Yidan Xu undertook the Mary Lister McCammon Summer Research Fellowship 2019 on related work.
abstract: |
  |  Spatial correlation between small-areas is typically accounted for using spatial random effects. However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions for non-grid geometries. Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous geostatistical data. Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
  |
  |  In a simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications. The simulations were performed under model misspecification and across a range of vignette and realistic geometries with different regularities. We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
date: "`r Sys.Date()`"
bibliography: citations.bib
#linenumbers: true
#numbersections: true
output: 
  bookdown::pdf_book: 
    base_format: rticles::isba_article
    md_extensions: -autolink_bare_uris #keep these off or \printead{} will fail on email addresses
    includes:
      in_header: preamble.tex
---

@@

Spatial random effects are frequently used to model spatial variation in areal data [@haining2003spatial; @cramb2018investigation].
The most common class of models used to specify spatial random effects are Gaussian Markov random fields (GMRFs), which combine a Gaussian distribution with Markov conditional independence assumptions between areas [@rue2005gaussian].
Under the Markov assumptions, observations made in areas close together are correlated, and more distant relationships are ignored.
Perhaps the simplest GMRF model is that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
This model is attractive as it requires minimal additional choices and is widely implemented.
As such, it has seen widespread use, including in agriculture [@oliver2015soil], ecology [@saracco2010modeling], epidemiology [@lawson2013statistical], image analysis [@schmid2006bayesian], neuroscience [@gossl2001bayesian] and public health [@dwyer2015drinking].
However, for irregular geometries like the administrative divisions of a country, the assumptions made by the Besag model about space are unrealistic.
As such, in this work we test the hypothesis that using more realistic assumptions about spatial structure improves the performance of small-area estimation models.
In doing so, we provide practical recommendations and further understanding of models for spatial structure.
<!-- @gutreuter2019improving use the SAR model, which is still adjacency based but is not exactly the Besag model--restructure introduction to be more general and include introduction of other adjacency based models (which all have the same criticisms)? -->

Our motivating application is the small-area estimation of HIV epidemic indicators in sub-Saharan Africa.
Estimates are required to help plan, implement, and evaluate the success of programmes, ensuring that available resources are most effectively used to respond to the epidemic [@hallett2016evaluation; @cuadros2017mapping].
Household surveys provide nationally represenative data about the general population, but are expensive to carry out and as such have small sample sizes at a district level.
Although auxiliary covariates can be used to aid estimation, data on the covariates most strongly associated with HIV, such as sexual risk behaviour [@howes2023spatio] or the prevalence of male circumcision, often face analogous measurement difficulties.
Possibly as a result, models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}.
This stands in contrast to other infectious diseases like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
These circumstances foreground the importance of models for spatial structure in HIV mapping.

Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
@duncan2017spatial compared seventeen methods for specifying these weights, but did not find any which outperform the Besag model.
This conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) which is recommended against by @vehtari2017practical.
Another approach is to take into account that areal data is typically produced by aggregating higher resolution data.
Obtaining inference about a variable at a different resolution to that it was observed at is known in geostatistics as the change-of-support problem [@gelfand2001change], and dates to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
@kelsall2002modeling applied ideas from change-of-support modelling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially].
<!-- `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages. -->
Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2016tmb], and made available as a part of the `disaggregation` package [@nandi2020disaggregation].

@@

Let $\Sc \subset \mathbb{R}^2$ be the study region, and the disjoint areas $\{A_i\}_{i = 1}^n$ be a spatial partition of $\Sc$ such that $A_i \cap A_j = \emptyset$ and $\cup_{i = 1}^n A_i = \Sc$.
Areal data are a type of spatial data where observations $y_i \in \mathcal{Y}$ are associated to areas $A_i$.
Examples of areal data include the colour of a pixel, the minimum wage in a state, and the number of disease cases in a region.
Geostatistical data are another type of spatial data where instead observations $y(s) \in \mathcal{Y}$ can be made at any location $s \in \Sc$ of a spatially-continuous stochastic process.
Examples of geostatistical spatial data include the temperature at a monitoring station, the response to a household survey, and the quality of a soil sample.

Areal data may often be conceptualised as arising from aggregation of geostatistical data, such that $y_i = \int_{A_i} y(s) \text{d}s$.
Notable exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures, which have a truly discrete spatial structure.
<!-- Note that, for data of this sort, there is less clear justification for using spatial smoothing. -->
<!-- Realistically, data relate to many covariates, some number of which are likely have discrete spatial structure and others of which continuous spatial structure, such that the result is often an intermediate. -->

In some situations it may be preferable to work using aggregated areal data rather than the original geostatistical data.
For one, the geostatistical data may be unavailabile or inaccessible due to privacy constraints, administrative practicality or limitations in storage capacity.
Alternatively, opting to model at the area-level can be advantageous as models for geostatistical data tend to be more complex, require more data, have higher computational burden, and offer less immediate applicablility to area-level policy-making.
In addition, in some circumstances, seemingly geostatistical data may truly be areal, such as data observed at polling stations or health facilities which individuals from the surrounding area travel to.
Ultimately, the decision to model the underlying data generating mechanism as either areal or geostatistical is often a matter of practical considerations and pragmatism.

@@

Auto-correlation is an important property of most spatial processes.
Usually, outcomes for locations close together in space tend to be more similar than those far apart, known as "Tobler's first law of geography" [@tobler1970computer].
This property both presents a challenge and an opportunity.
Each observation provides less information than it would have had the samples been independent, making it more difficult to estimate global parameters.
On the other hand, spatial correlation can be used to improve local parameter estimates, indexed by a particular spatial location, especially in parts of the study region where little to no information is available.

The latter benefit is basis for the statistical task of small-area estimation, which aims to produce reliable local estimates where small sample sizes lead to noisy data [@pfeffermann2013new].
In a spatial setting, this is often in small geographic areas. 
However, the phrase "small-area" is not restricted to geographic areas and may be interpreted more broadly to mean any area where data are insufficient to make accurate local parameter inferences.
For example, in the context of multilevel regression and post-stratification [@gelman1997poststratification], small-areas are generated by the intersection of demographic variables like age, gender and race, alongside geographic variables.
Because of the cost of gathering samples, a survey may be designed to give reliable estimates at at an aggregate level but not at a small-area level.
Although direct estimators of local parameters are unbiased, when data are sparse the total error may be reduced by accepting some bias in exchange for reduced variance using so-called indirect estimators.
Smoothing approaches use information from similar units to "borrow strength" from one parameter to another, with determining precisely what is meant by "similar" a central challenge.
See @wakefield2020small for a recent review of both design-based and model-based approaches to small-area estimation of health-indicators, including both area and unit-level analysis.

@@

Bayesian hierarchical models are an attractive framework for small-area estimation.
Let $[n]$ denote $1, \ldots, n$.
Areal data $\y = (y_i)_{i \in [n]}$ may be modelled using a three-layer structure [@berliner1996hierarchical; @cressie2015statistics; @rao2015small] given by
\begin{alignat}{2}
&\text{(Observations)}  &        \y &\sim p(\y \, | \, \x, \btheta_1), \label{eq:data} \\
&\text{(Latent field)}  &        \x &\sim p(\x \, | \, \btheta_2), \label{eq:process} \\
&\text{(Parameters)}    & \qquad (\btheta_1, \btheta_2) &= \btheta \sim p(\btheta), \label{eq:parameters}
\end{alignat}
where $\x = (x_i)_{i \in [N]}$ is a $N$-dimensional latent field, and $\btheta = (\theta_i)_{i \in [m]}$ are the $m$-dimensional hyperparameters, with $m < N$.
Small-area estimation models are usually descriptive rather than mechanistic, so a natural choice is to model the latent field with a flexible and computationally efficient Gaussian distribution $\x \sim \mathcal{N}(\bmu, \mathbf{Q})$ with mean vector $\bmu$ and precision matrix $\mathbf{Q}$.
Three-layer models with Gaussian latent fields have been collectively studied as latent Gaussian models (LGMs) [@rue2009approximate] and comprise a wide array of popular models, including generalised linear mixed models.

@@

Spatial random effects $\phi(\cdot)$ are one Gaussian function which may be included in the model, whose role is to capture the spatial structure between areas not already accounted for by $\beta_0 + \sum_{l = 1}^{p} \beta_j z_{ji}$.
Should no spatial structure remain, then independent and identically distributed (IID) random effects $\phi_i \sim \N (0, \tau_\phi^{-1})$ may be appropriate, where $\tau_\phi$ is a shared precision.
Exploratory data analysis techniques such as visual inspection or Moran's $I$ coefficient [@cliff1981spatial] may be used to determine if there remains any spatial auto-correlation in the data.
<!-- TODO: Add calculation of Moran's $I$ to my HIV survey data-sets. -->
Specifying $\bphi = (\phi_i)_{i \in [n]}$ amounts to specifying the entries of a precision or covariance matrix, as we discuss in the following sections.

@@

The Besag model was originally proposed for use in image analysis, where areas correspond to pixels arranged in a regular lattice structure.
Since then, it has seen wider use, including in situations where the spatial structure is less regular.
We highlight a number of concerns about the model's applicability to this broader setting.
This discussion is closely linked to the modifiable areal unit problem [@openshow1979million], whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, as well as the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation].
<!-- Missing an item about potential loss of information by using a sparse matrix, though I don't actually think this is very important. -->

@@

Summarising a geometry by an adjacency graph represents a loss of information.
Many geometries share the same adjacency graph, and therefore the same probability model, as illustrated by Figure \ref{fig:tikz1}.
This fact is not concerning in itself, but does prompt consideration as to whether the class of geometries with the same adjacency graph is sufficiently similar to merit identical models.^[The regularity of realistic geometries may help to constrain each class to be more similar than it strictly has to be. In other words, although pathological geometries can be constructed, they are implausible in statistical practise and so not of great concern to us here.]
Intuitively the more regular the spatial structure, the less information is lost in compression to an adjacency graph.
In image analysis, very little spatial information is lost in compression of a lattice structure to an adjacency graph.
On the other hand, the regions of a country, determined by political and geographic forces, tend to display greater irregularity.
As such, the appropriateness of adjacency compression varies by the type of geometry common to the application setting.

@@

In Equation \ref{eq:besag} the precision of $\phi_i$ is proportional to its number of neighbours $n_{\delta i}$.
It follow that as $n_{\delta i} \to \infty$ then $\text{Var}(\phi_i) \to 0$.
This is illustrated by Figure \ref{fig:tikz2} where the area on the right is repeatedly divided such that its number of neighbours increases.
This property is a consequence of averaging the conditional mean over a greater number of areas, which, in certain situations, can correspond to a greater amount of information.
However, if the amount of information in the shaded area remains fixed, it seems inappropriate that $\text{Var}(\phi_1)$ should tend to zero simply as a result of drawing additional, arbitrary, boundaries.
In the image analysis setting this modelling assumption is reasonable: each pixel represents a fixed amount of information and a higher pixel density represents a greater amount of information.
On the other hand, in public health and epidemiology, drawing boundaries to create additional areas is not expected to correspond to a greater amount of information.

@@

The simplest approach is to use a single point.
A natural choice is the centroid $c_i \in A_i$, given by the arithmetic mean of the latitude and longitude, which may be representative of the area.^[Though note that it is not guaranteed for the centroid to (even) lie within the area.]
This results in the centroid kernel
\begin{equation}
    K(A_i, A_j) = k(c_i, c_j).
\end{equation}
<!-- More broadly, given a weighting distribution $W_i$ with probability density proportional to $w_i(s)$ over $A_i$, weighted centroids could be used in the above equation. -->
The centroid kernel has been used in environmental epidemiology [@wakefield1999spatial] and to model the reproduction number of COVID-19 [@teh2021efficient].
In a model comparison study \citet[Section 3]{best2005comparison} simulated data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas.^[Though it is unsuprising that a stationary covariance would struggle to recover non-stationary structure.]

@@

Rather than looking to choose a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest.
This results in the integrated kernel
\begin{equation}
K(A_i, A_j) = \frac{1}{|A_i||A_j|} \int_{A_i} \int_{A_j} k(s, s') \text{d} s \text{d} s'. \label{eq:ik1}
\end{equation}
This covariance structure is equivalent to that obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition, and has been studied in the machine learning literature under the title aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi].
<!-- @law2018variational develop a variational Bayesian inference method for aggregated observation models. -->
<!-- @tanaka2019spatially consider aggregation over a linear combination of latent Gaussian processes, each of which is associated to data corresponding to a different partition of the study region. -->
Unlike for the centroid kernel where $K_{ii} = 1$ for all $i$, the marginal variance of the $i$th spatial random effect $K_{ii} = K(A_i, A_i)$ varies depending on the area: becoming smaller for more compact areas and larger for areas which are of greater extent or more spread out.

@@

Additional information accounting for heterogeneity over $A_i$ may be incorporated into the integrated kernel.
This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure, to give a weighted integrated kernel
\begin{equation}
K(A_i, A_j) = \frac{1}{|A_i||A_j|}\int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d} s \text{d} s', \label{eq:ik2}
\end{equation}
This may be useful in disease mapping, where we expect regions with populations who live close to their shared border to be more strongly correlated than regions whose populations live far apart, which could be accounted for by weighting according to a high resolution measure of population density.

@@

Most of the time we do not expect to be able to calculate Equation \ref{eq:ik2} analytically.
Instead, given $n$ collections of $L_i$ samples $\{s^{(i)}_l \}_{l = 1}^{L_i} \sim \mathcal{U}(A_i)$ drawn uniformly from each area then the integral may be approximated using Monte Carlo by the double sum
\begin{equation}
K(A_i, A_j) \approx \frac{1}{L_i L_j} \sum_{l = 1}^{L_i} \sum_{m = 1}^{L_j}
w_i \left( s^{(i)}_l \right) w_j \left( s^{(j)}_m \right)
k \left( s^{(i)}_l, s^{(j)}_m \right). \label{eq:mcintegratedkernel}
\end{equation}
Equivalently, samples drawn from $W_i$ may be used without weighting by $w_i(s)$.
Nodes may also be selected by some deterministic process to give a numerical quadrature estimate of the kernel.
These approaches require $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $K$.
This imposes a significant computational cost if the Gram matrix is often recomputed during inference, as is the case in MCMC when any of the kernel hyperparameters are learnt, placing a limit on the number of samples or nodes it is feasible to use.
@kelsall2002modeling make inference more feasible by using a discrete hyperparameter prior to reduce the number of Gram matrix constructions and inversions required.

>>>>

---
title: |
  | Beyond borders: evaluating the suitability of spatial adjacency for small-area estimation
classoption: ba
author:
  - firstname: Adam 
    lastname: Howes
    email: ath19@ic.ac.uk
    affiliationref: A,B
    footnoterefs: 
      - t1
  - firstname: Jeffrey W.
    lastname: Eaton
    affiliationref: C,B
  - firstname: Seth R.
    lastname: Flaxman
    affiliationref: D
affiliations:
  - ref: A
    name: Imperial College London 
    address: Department of Mathematics
  - ref: B
    name: Imperial College London
    address: MRC Centre for Global Infectious Disease Analysis
  - ref: C
    name: Harvard University
    address: Center for Communicable Disease Dynamics, Harvard T.H. Chan School of Public Health
  - ref: D
    name: University of Oxford
    address: Department of Computer Science 
footnotes:
  - ref: t1
    text: "Corresponding Author"
supplements:
  - title: Appendix
    description: All further materials for this study.
acknowledgements: |
  | AH thanks Robert Ashton for help with \textsf{R} package development, and Harrison Zhu, Theo Rashid, Tim Wolock, Samir Bhatt, Tim Lucas, Elizaveta Semenova, Marta Blangiardo and Oliver Ratmann for helpful conversations. Yidan Xu undertook the Mary Lister McCammon Summer Research Fellowship 2019 on related work.
abstract: |
  |  Spatial correlation between small-areas is typically accounted for using spatial random effects. However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic spatial assumptions for geometries which are not a grid. Furthermore, they do not take into account areal data is commonly generated by aggregating continuous geostatistical data. Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
  |
  |  In a simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications. The simulations were performed under model misspecification and across a range of vignette and realistic geometries with different regularities. We also compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
  |
  |  The results of this analysis are forthcoming, and will be described here.
date: "`r Sys.Date()`"
bibliography: citations.bib
#linenumbers: true
#numbersections: true
output: 
  bookdown::pdf_book: 
    base_format: rticles::isba_article
    md_extensions: -autolink_bare_uris #keep these off or \printead{} will fail on email addresses
    includes:
      in_header: preamble.tex
---

@@

Spatial random effects are frequently used to model spatial variation in areal data [@haining2003spatial; @cramb2018investigation].
The most common class of models used to specify spatial random effects are Gaussian Markov random fields (GMRFs).
These models combine a Gaussian distribution with Markov conditional independence assumptions between areas [@rue2005gaussian].
It is assumed that observations made in areas close together are correlated, and more distant relationships are ignored.
Perhaps the simplest GMRF model is that of @besag1991bayesian, where information is borrowed equally from each adjacent area.
This model is attractive as it requires minimal additional modelling choices and is widely implemented.
As such, it has seen widespread use, including in agriculture [@oliver2015soil], ecology [@saracco2010modeling], epidemiology [@lawson2013statistical], image analysis [@schmid2006bayesian], neuroscience [@gossl2001bayesian] and public health [@dwyer2015drinking].
However, for irregular geometries like the administrative divisions of a country, the assumptions made by the Besag model about space are unrealistic.
We test the hypothesis that using more realistic assumptions about spatial structure improves the performance of small-area estimation models.
In doing so, we provide practical recommendations and further understanding of models for spatial structure.
<!-- @gutreuter2019improving use the SAR model, which is still adjacency based but is not exactly the Besag model--restructure introduction to be more general and include introduction of other adjacency based models (which all have the same criticisms)? -->

Our motivating application is the small-area estimation of HIV epidemic indicators in sub-Saharan Africa.
These estimates are required to help plan, implement, and evaluate the success of programmes, ensuring that available resources are most effectively used to respond to the epidemic [@hallett2016evaluation; @cuadros2017mapping].
Household surveys provide nationally represenative data about the general population, but are expensive to carry out, and as such have small sample sizes at a district level.
Although auxiliary covariates can be used to aid estimation, data on the covariates most strongly associated with HIV, such as sexual risk behaviour [@howes2023spatio] or the prevalence of male circumcision, often face similar measurement difficulties.
Potentially as a result, models including covariates have only been found to modestly improve predictions \citep[Supplementary Figure 20]{dwyer2019mapping}.
In contrast, covariates do see substantial use for mapping other infectious diseases, like Malaria where transmission is driven by more predictive and easily-measurable environmental factors [@weiss2015re].
These circumstances foreground the importance of models for spatial structure in HIV mapping.

Within the GRMF framework, attempts to more accurately reflect irregular spatial structure have defined weights specifying the extent to which neighbours are related.
@duncan2017spatial compared seventeen methods for specifying these weights, but did not find any which outperform the Besag model.
This conclusion was specific to the often-analysed Scottish lip cancer example, and based on the deviance information criteria (DIC) recommended against by @vehtari2017practical.
Another approach is to take into account that areal data is typically produced by aggregating higher resolution data.
Obtaining inference about a variable at a different resolution to that it was observed at is known in geostatistics as the change-of-support problem [@gelfand2001change], and dates to foundational work on block kriging [@krige1951statistical; @matheron1976forecasting].
@kelsall2002modeling applied ideas from change-of-support modelling to consider areal data as coming about by an aggregation of a continuously-indexed Gaussian process, resulting in a covariance structure between two areas given by the average covariance between two points chosen randomly from those areas.
This type of model is particularly natural in the log-Gaussian Cox process modelling framework [@li2012log; @diggle2013spatial; @taylor2015bayesian; @johnson2019spatially].
<!-- `lgcp` [@taylor2015bayesian] and `SDALGCP` [@johnson2019spatially] packages. -->
Inference for aggregated data has recently been advanced by @wilson2018pointless who consider the SPDE approach of @lindgren2011explicit, using `R-INLA`, and an empirical Bayes approach, using `TMB` [@kristensen2016tmb], and made available as a part of the `disaggregation` package [@nandi2023disaggregation].

@@

Let $\Sc \subset \mathbb{R}^2$ be the study region, and the disjoint areas $\{A_i\}_{i = 1}^n$ be a spatial partition of $\Sc$ such that $A_i \cap A_j = \emptyset$ and $\cup_{i = 1}^n A_i = \Sc$.
Areal data are a type of spatial data where observations $y_i \in \mathcal{Y}$ are associated to areas $A_i$.
Examples of areal data include the colour of a pixel, the minimum wage in a state, and the number of disease cases in a region.
Geostatistical data are another type of spatial data where instead observations $y(s) \in \mathcal{Y}$ can be made at any location $s \in \Sc$ of a spatially-continuous stochastic process.
Examples of geostatistical spatial data include the temperature at a monitoring station, the response to a household survey, and the quality of a soil sample.
Areal data may often be conceptualised as arising from aggregation of geostatistical data, such that $y_i = \int_{A_i} y(s) \text{d}s$.
Notable exceptions include policies determined at an administrative level, such as the minimum wage or disease control measures.
<!-- Note that, for data of this sort, there is less clear justification for using spatial smoothing. -->
<!-- Realistically, data relate to many covariates, some number of which are likely have discrete spatial structure and others of which continuous spatial structure, such that the result is often an intermediate. -->

In some situations it may be preferable to work using aggregated areal data rather than the underlying geostatistical data.
For one, the geostatistical data may be unavailabile or inaccessible due to privacy constraints, administrative practicality or limitations in storage capacity.
Alternatively, opting to model at the area-level can be advantageous: models for geostatistical data tend to be more complex, require more data, have higher computational burden, and offer less immediate applicability to policy.
Under some circumstances, seemingly geostatistical data may truly be areal, as with data observed at polling stations or health facilities which individuals from the surrounding area travel to.
Ultimately, the decision to model the data generating mechanism as either areal or geostatistical is often a matter of practical considerations and pragmatism.
This is reflected in the variety of approaches taken by modellers.

@@

Auto-correlation is an important property of most spatial processes.
Usually, outcomes for locations close together in space tend to be more similar than those far apart, known as "Tobler's first law of geography" [@tobler1970computer].
This property presents both a challenge and an opportunity.
Each observation provides less information than it would have had the samples been independent, making it more difficult to estimate global parameters.
On the other hand, spatial correlation can be used to improve local parameter estimates, indexed by a particular spatial location, especially in parts of the study region where little to no information is available.

The latter benefit is basis for the statistical task of small-area estimation, which aims to produce reliable local estimates where small sample sizes lead to noisy data [@pfeffermann2013new].
In a spatial setting, this is often in small geographic areas. 
However, the phrase "small-area" is not restricted to geographic areas and may be interpreted more broadly to mean any area where data are insufficient to make accurate local parameter inferences.
For example, in the context of multilevel regression and post-stratification [@gelman1997poststratification], small-areas arise by the intersection of demographic variables like age, gender and race, alongside geographic variables.
Because of the cost of gathering samples, a survey may be designed to give reliable estimates at at an aggregate level but not at a small-area level.
Although direct estimators of local parameters are unbiased, when data are sparse the total error may be reduced by accepting some bias in exchange for reduced variance using so-called indirect estimators.
Smoothing approaches use information from similar units to "borrow strength" from one parameter to another. 
Determining precisely what is meant by "similar" a central challenge.
See @wakefield2020small for a recent review of both design-based and model-based approaches to small-area estimation of health-indicators, including both area and unit-level analysis.

@@

Bayesian hierarchical models are an attractive framework for small-area estimation.
Let $[n]$ denote $1, \ldots, n$.
Areal data $\y = (y_i)_{i \in [n]}$ may be modelled using a three-layer structure [@berliner1996hierarchical; @cressie2015statistics; @rao2015small] given by
\begin{alignat}{2}
&\text{(Observations)}  &        \y &\sim p(\y \, | \, \x, \btheta_1), \label{eq:data} \\
&\text{(Latent field)}  &        \x &\sim p(\x \, | \, \btheta_2), \label{eq:process} \\
&\text{(Parameters)}    & \qquad (\btheta_1, \btheta_2) &= \btheta \sim p(\btheta), \label{eq:parameters}
\end{alignat}
where $\x = (x_i)_{i \in [N]}$ is a $N$-dimensional latent field, and $\btheta = (\theta_i)_{i \in [m]}$ are the $m$-dimensional hyperparameters, with $m < N$.
Small-area estimation models are usually descriptive rather than mechanistic, so a natural choice is to model the large latent field with a flexible and computationally efficient Gaussian distribution $\x \sim \mathcal{N}(\bmu, \mathbf{Q})$ with mean vector $\bmu$ and precision matrix $\mathbf{Q}$.
Three-layer models with Gaussian latent fields have been collectively studied as latent Gaussian models (LGMs) [@rue2009approximate] and comprise a wide array of popular models, including generalised linear mixed models.

@@

Spatial random effects $\phi(\cdot)$ are a Gaussian function, whose role is to capture the spatial structure between areas not already accounted for by the intercept and linear effects of the covariates $\beta_0 + \sum_{l = 1}^{p} \beta_j z_{ji}$.
Should no spatial structure remain, then independent and identically distributed (IID) random effects $\phi_i \sim \N (0, \tau_\phi^{-1})$ may be appropriate, where $\tau_\phi$ is a shared precision.
Exploratory data analysis techniques such as visual inspection or Moran's $I$ coefficient [@cliff1981spatial] may be used to determine if there remains any spatial auto-correlation in the data.
<!-- TODO: Add calculation of Moran's $I$ to my HIV survey data-sets. -->
Specifying $\bphi = (\phi_i)_{i \in [n]}$ amounts to specifying the entries of a precision or covariance matrix, as we discuss in the following sections.

@@

The Besag model was originally proposed for use in image analysis, where areas correspond to pixels arranged in a regular lattice structure.
Since then, it has seen wider use, including in situations, like our work mapping HIV, where the spatial structure is less regular.
We have a number of concerns about the model's applicability to this broader setting.
This discussion is closely linked to the modifiable areal unit problem [@openshow1979million], whereby statistical conclusions change as a result of seemingly arbitrary changes in data aggregation, as well as the challenge of ecological inference and the ecological fallacy [@wakefield2010aggregation].
<!-- Missing an item about potential loss of information by using a sparse matrix, though I don't actually think this is very important. -->

@@

Summarising a geometry by an adjacency graph represents a loss of information.
Many geometries share the same adjacency graph, and are as such isomorphic identical under the Besag model (Figure \ref{fig:tikz1}).
This is not in itself a problem, but does prompt consideration as to whether the class of geometries with the same adjacency graph is sufficiently similar to merit identical models.^[The regularity of realistic geometries may help to constrain each class to be more similar than it strictly has to be. In other words, although pathological geometries can be constructed, they are implausible in statistical practise and so not of great concern to us here.]
Intuitively, the more regular the spatial structure, the less information is lost in compression to an adjacency graph.
In image analysis, very little spatial information is lost in compression of a lattice structure to an adjacency graph.
On the other hand, the regions of a country, determined by political and geographic forces, tend to display greater irregularity.
The appropriateness of adjacency compression therefore varies by the type of geometry common to the application setting.

@@

In Equation \ref{eq:besag} the precision of $\phi_i$ is proportional to its number of neighbours $n_{\delta i}$.
It follow that as $n_{\delta i} \to \infty$ then $\text{Var}(\phi_i) \to 0$.
This is illustrated by Figure \ref{fig:tikz2} where the area on the right is repeatedly divided such that its number of neighbours increases.
This property is a consequence of averaging the conditional mean over a greater number of areas, which, in certain situations, can correspond to a greater amount of information.
However, if the amount of information in the shaded area remains fixed, it is inappropriate that $\text{Var}(\phi_1)$ should tend to zero as a result of drawing additional, arbitrary, boundaries.
In the image analysis setting this modelling assumption is reasonable: each pixel represents a fixed amount of information and a higher pixel density represents a greater amount of information.
On the other hand, in public health and epidemiology, drawing boundaries to create additional areas is not expected to correspond to a greater amount of information.

@@

The simplest approach is to use a single point such that $K(A_i, A_j) = k(p_i, p_j)$.
A natural choice is the centroid $p_i = c_i$, given by the arithmetic mean of the latitude and longitude, which may be representative of the area.^[Note it is not guaranteed for the centroid to (even) lie within the area i.e. we may have $c_i \notin A_i$.]
This results in the centroid kernel
\begin{equation}
    K(A_i, A_j) = k(c_i, c_j).
\end{equation}
The centroid kernel has been used in environmental epidemiology [@wakefield1999spatial] and to model the reproduction number of COVID-19 [@teh2021efficient].
<!-- In a model comparison study \citet[Section 3]{best2005comparison} simulated data representing heterogeneous exposure to air pollution, including elevated rates of exposure near two hypothetical point source locations, and found that the centroid kernel tended to over-smooth the high-risk areas.^[Though it is unsuprising that a stationary covariance would struggle to recover non-stationary structure.] -->

@@

Rather than choosing a single representative point, an alternative is to represent the whole area by integrating the kernel over the areas of interest.
This results in the integrated kernel
\begin{equation}
K(A_i, A_j) = \frac{1}{|A_i||A_j|} \int_{A_i} \int_{A_j} k(s, s') \text{d} s \text{d} s'. \label{eq:ik1}
\end{equation}
This covariance structure is equivalent to that obtained by aggregating a spatially continuous Gaussian process with kernel $k$ over the areal partition, and has been studied in the machine learning literature under the name aggregated Gaussian processes [@law2018variational; @tanaka2019spatially; @yousefi2019multi; @hamelijnck2019multi].
<!-- @law2018variational develop a variational Bayesian inference method for aggregated observation models. -->
<!-- @tanaka2019spatially consider aggregation over a linear combination of latent Gaussian processes, each of which is associated to data corresponding to a different partition of the study region. -->
Unlike for the centroid kernel where $K_{ii} = 1$ for all $i$, the marginal variance of the $i$th spatial random effect $K_{ii} = K(A_i, A_i)$ varies depending on the area: becoming smaller for more compact areas and larger for areas which are of greater extent or more spread out.

@@

Additional information accounting for heterogeneity over $A_i$ may be incorporated into the integrated kernel^[Analogously, weighted centroids could also be used in the centroid kernel.]
This can be accomplished using weighting distributions $\{W_i\}$ which represent an unequal contribution of each point to the similarity measure, to give a weighted integrated kernel
\begin{equation}
K(A_i, A_j) = \frac{1}{|A_i||A_j|}\int_{A_i} \int_{A_j} w_i(s) w_j(s') k(s, s') \text{d} s \text{d} s', \label{eq:ik2}
\end{equation}
This may be useful in disease mapping, where we expect regions with populations who live close to their shared border to be more strongly correlated than regions whose populations live far apart, which could be accounted for by weighting according to a high resolution measure of population density.

@@

Most of the time we do not expect to be able to calculate Equation \ref{eq:ik2} analytically.
Instead, given $n$ collections of $L_i$ samples $\{s^{(i)}_l \}_{l = 1}^{L_i} \sim \mathcal{U}(A_i)$ drawn uniformly from each area then the integral may be approximated using Monte Carlo by the double sum
\begin{equation}
K(A_i, A_j) \approx \frac{1}{L_i L_j} \sum_{l = 1}^{L_i} \sum_{m = 1}^{L_j}
w_i \left( s^{(i)}_l \right) w_j \left( s^{(j)}_m \right)
k \left( s^{(i)}_l, s^{(j)}_m \right). \label{eq:mcintegratedkernel}
\end{equation}
Equivalently, samples drawn from $W_i$ may be used without weighting by $w_i(s)$.
Nodes may also be selected deterministically to give a numerical quadrature estimate of the kernel.
These approaches require $\mathcal{O}(\sum_{i = 1}^n \sum_{j = 1}^n L_i L_j)$ evaluations of the kernel $k$ to compute the $n \times n$ Gram matrix $K$.
This imposes a significant computational cost if the Gram matrix is often recomputed during inference, as is the case in MCMC when any of the kernel hyperparameters are learnt, placing a limit on the number of samples or nodes it is feasible to use.
@kelsall2002modeling make inference more feasible by using a discrete hyperparameter prior to reduce the number of Gram matrix constructions and inversions required.
